Update of the ozone dry deposition in the OsloCTM3 v1.0

### 2018-11-27 DryDep.tex
Wrote the conclusions, but not sure what should go to discussions... corrected some typos.

In Section~\ref{sec:model_des}, we will briefly describe the OsloCTM3, give a detailed account of the new dry deposition scheme (Section~\ref{subsec:DryDep}) as well as present pre-processing of meteorological input data to compute necessary input to the dry deposition scheme such as begin and duration of greening season (GDAY, GLEN) and photosynthetic photon flux density (PPFD) (Section~\ref{subsec:pre-pro}). In Section~\ref{sec:eval}, we present sensitivity tests with respect to a manifold of parameters in the dry deposition scheme (Section~\ref{subsec:sens}) and validate our results with respect to results from the multi-model intercomparison of \citet{ACP:Hardacre2015} (Section~\ref{subsec:model}) and to observations both at the surface as well as in the free troposphere (Section~\ref{subsec:obs}). In Section~\ref{sec:disc}, we will summarize our results and draw conclusions for further development of the model (Section~\ref{sec:conc}).
{\bf TODO: Precise outlook on paper when writing is done.}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Model description}
\label{sec:model_des}
The OsloCTM3 is a three dimensional global chemistry transport model (CTM). The key components of the OsloCTM3 have been described and evaluated by \citet{GMD:Sovde2012}. A detailed account of the capabilities of the OsloCTM3 in simulating anthropogenic aerosol forcing in the past and recent past using the Community Emission Data System (CEDS) historical emission inventory \citep{GMD:Hoesly2018} is given by \citet{GMD:Lund2018}. The OsloCTM3 can also be coupled to the Model of Emissions of Gases and Aerosols from Nature (MEGAN v2.10) \citep{ACP:Guenther2006}. A publication focusing on this is planed. Parallelization of the OsloCTM3 has not been optimized as of now. While the meteorological data driving the OsloCTM3 is given in a resolution of at T159N80L60, it is very time and memory consuming to run the CTM with full chemistry at this resolution. Therefore, we run with a reduced horizontal resolution of $2.25^\circ\times2.25^\circ$. In the following, we will give a detailed account of the new dry deposition scheme. Although some of the equations in this section may be \emph{textbook knowledge}, for the sake of completeness, it is important to summarize them, nonetheless.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Ozone dry deposition scheme}
\label{subsec:DryDep}
In the original dry deposition scheme of the OsloCTM3, the state of the atmosphere was not taken into account. Dry deposition velocities were rather parameterized following the work of \citet{AE:Wesely1989} with parameter updates from \citet{JGR:Hough1991}. This means that seasonal day and night average deposition velocities for different land surface types (water, forest, grass, tundra/desert, and ice and snow) were in use. Day was distinguished from night by solar zenith angles below $90\,\unit{^\circ}$. Winter was defined by temperatures below $273.15\,\unit{K}$ for gridboxes containing land masses. For ocean, winter and summer parameters are equal in this parameterization, therefore no distinctive treatment is needed for ocean gridboxes. In addition, a reduced uptake due to snow cover above $1\,\unit{m}$ for forest and $10\,\unit{cm}$ for grass/tundra, respectively, was taken into account. We will refer to this parameterization as \emph{Wesely scheme}.\\

@@

\subsubsection*{Mapping of land surface types}
Land surface types from, e.g., \href{https://daac.ornl.gov/cgi-bin/dataset_lister.pl?p=29}{ISLSCP2} product from MODIS or Community Land Model (\href{http://www.cgd.ucar.edu/tss/clm/}{CLM}) 2 categories, have to be mapped to the nine land surface types used in the EMEP scheme (Fig.~\ref{fig:pft_mapping}). For both, MODIS and CLM2 land surface categories, snow and ice cover is estimated from input meteorology, while water is defined as $1-\sum_{N=0}^{N_\text{max}} f_L(N)$. From the MODIS category \emph{Barren or sparsely vegetated}, everything polward from $60\,\unit{^\circ}$ is defined as tundra, while everything equatorward is categorized as desert. This mapping differs from the one used in the Wesely scheme.

@@

  \caption{Mapping of land surface categories from ISLSCP2 product of MODIS /  Community Land Model (CLM) 2 to the nine land surface types used in the EMAP scheme. Water bodies and snow and ice categories of MODIS are actually not mapped. For both, MODIS and CLM2 land surface categories, snow and ice cover is estimated from input meteorology, while water is defined as $1-\sum_{N=0}^{N_\text{max}} f_L(N)$. From MODIS category 'Barren or sparsely vegetated', everything polward from $60\,\unit{^\circ}$ is defined as tundra, while everything equatorward is categorized as desert.}

@@

The annual total ozone dry deposition is shown in Fig.~\ref{fig:mmm_drydep}c). In accordance to the previously described features, we observe a reduction of the global total ozone dry deposition in all sensitivity studies to almost one half of the amount given by \emph{Wesely\_type}. The occurrence of this reduction in the zonal bands where the major deserts are located points to a change in $v_\text{desert}^\chem{O_3}$. Consulting the parameter file used in the Wesely scheme, we find $v_\text{desert}^\chem{O_3} \equiv v_\text{tundra}^\chem{O_3} = 0.26\,\unit{cm\,s^{-1}}$ \citep{JGR:Hough1991}, while in the EMEP scheme $v_\text{desert}^\chem{O_3}=0.05\,\unit{cm\,s^{-1}}$ and $v^\chem{O_3}_\text{tundra} = 0.24\,\unit{cm\,s^{-1}}$, respectively. Similarly, dry deposition velocities over ice and snow and ocean have been even higher in the Wesely scheme ($v_\text{ice/snow}^\chem{O_3}\equiv v_\text{water}^\chem{O_3}=0.07\,\unit{cm\,s^{-1}}$) than in the original EMEP scheme ($v_\text{ice/snow}^\chem{O_3}\equiv v_\text{water}^\chem{O_3}=0.05\,\unit{cm\,s^{-1}}$). These differences in baseline dry deposition velocities over huge parts of the unvegetated surface of the Earth amount to most of the quantitative difference between the Wesely and the EMEP scheme. We further elaborate on this in the following section (Section~\ref{subsubsec:seasons}).

\begin{figure}[t]
  \includegraphics[width=8.3cm]{pictures/final-vmr_surfaceozone}\\
  \includegraphics[width=8.3cm]{pictures/final-total_ozone_drydepvelo}\\
  \includegraphics[width=8.3cm]{pictures/final-total_ozone_drydep}
  \caption{Comparison of the manifold of OsloCTM3 integrations with respect to (a) Surface ozone concentrations, (b) Annual average ozone dry deposition velocity, (c) Total annual ozone dry deposition. The different colors indicate sets of simulation with similar baselines. The multi-model mean from the evaluation of HTAP models by \citet{ACP:Hardacre2015} is shown as a reference (where available).}
  \label{fig:mmm_drydep}
\end{figure}
%
In Fig.~\ref{fig:mmm_drydep_season}a), the average zonal ozone dry deposition is shown separated by month. Where available, we have added the multi-model-mean given by \citet{ACP:Hardacre2015} as reference. There are two major phases apparent: NH and SH greening season. Former is reflected in a pronounced peak of $v_\text{DD}^\chem{O_3}$ in the northern mid latitudes, while the absence of this peak can be attributed to the latter in addition to a southward shift of the tropical peak dry deposition velocity. In the Wesely scheme, NH mid latitude peak velocities appear in June compared to July in the EMEP scheme, indicating that the prescribed seasonal cycles differ. The corresponding total monthly ozone dry deposition is shown in Fig.~\ref{fig:mmm_drydep_season}b). In general, the seasonal patterns are quite similar in the Wesely scheme and the EMEP scheme, displaying a strong symmetry around $10\,\unit{^\circ N}$ in January/February and November/December, respectively. What differs most is the molding and intensity of the NH peak dry deposition. Both schemes reach the maximum in June/July but the peak is much more differentiated in March already in the Wesely scheme. Similarly, the SH tropical peak dry deposition is reached in August/September but sustained longer, into October, in the Wesely scheme. Since we have not conducted any simulation with a meteorological year other than 2005, we cannot elaborate on whether this is a special feature of our chosen year or not. In comparison, the available two month of the multi-model-mean display a different pattern. Especially the NH/SH asymmetry is much more pronounced in February, than it is the EMEP scheme.
%
\begin{figure}[t]
  \includegraphics[width=8.3cm]{pictures/final-total_ozone_drydepvelo_monthly}\\
  \includegraphics[width=8.3cm]{pictures/final-total_ozone_drydep_monthly}
  \caption{Comparison of the manifold of OsloCTM3 integrations with respect to (a) Zonal average ozone dry deposition velocities; (b) Total annual amount of ozone removed from the atmosphere via dry deposition. The multi-model mean from the evaluation of HTAP models by \citet{ACP:Hardacre2015} is shown as a reference (where available).}
  \label{fig:mmm_drydep_season}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Average seasonal cycles}
\label{subsubsec:seasons}
%
To further disentangle the contributions of different regions to the global ozone budget, we will briefly look at different projections of seasonal cycles.

@@

As suggested by \citet{ACP:Hardacre2015}, we retrospectively separate ozone dry deposition velocities with respect to surface types. Based on a CLM~2 average dynamic land surface map, we generated masks for 9 different surface types (Fig.~\ref{fig:pft_landsurface}a) and used these to select gridboxes with a high percentage of these surface types, ranging from a meager $70\,\unit{\%}$ for cropland in the NH mid latitudes to $100\,\unit{\%}$ regarding desert, ocean, snow and ice, and tropical forest. Thus, it was not possible to exclusively select gridboxes with $100\,\unit{\%}$ cover for every surface type. Since we have not performed a full unfolding on the data, the results should be treated with slight caution. In Fig.~\ref{fig:mmm_drydep_season_pft}, the seasonal cycles of dry deposition velocities are shown for the nine surface categories. The patterns and absolute numbers differ substantially between the Wesely scheme and the EMEP scheme and the multi-model-mean. The divergence of the average dry deposition velocities between the Wesely scheme and the EMEP scheme in desert regions ($\Delta \overline{v}_\text{desert}^\chem{O_3} = 0.14\,\unit{cm\,s^{-1}}$) as well as grassland ($\Delta \overline{v}_\text{desert}^\chem{O_3} = 0.4\,\unit{cm\,s^{-1}}$) is quite remarkable as well as the difference of both to the multi-model-mean in tropical forest regions ($\Delta \overline{v}_\text{desert}^\chem{O_3} = 0.38\,\unit{cm\,s^{-1}}$). In contrast to the Wesely scheme and the EMEP scheme, the multi-model-mean indicates a rather pronounced seasonal cycle in desert regions ($0.10\,\unit{cm\,s^{-1}} \leq v^\chem{O_3}_\text{desert}\leq 0.15\,\unit{cm\,s^{-1}}$). The estimated dry deposition velocities over desert regions are consistent with the previously mentioned baseline values of ozone dry deposition, which means that in the EMEP scheme they are about one order of magnitude lower than in the Wesely scheme. From measurements of ozone fluxes in the Sahara desert, \citet{AE:Gusten1995} deduced $v^\chem{O_3}_\text{desert, day} = 0.1\,\unit{cm\,s^{-1}}$, $v^\chem{O_3}_\text{desert, night} = 0.04\,\unit{cm\,s^{-1}}$, and $\overline{v}^\chem{O_3}_\text{desert} = 0.065\,\unit{cm\,s^{-1}}$. This implies, that ozone dry deposition over desert regions is highly overestimated in the Wesely scheme as well as in the multi-model-mean, while it may be underestimated in the EMEP scheme. Similarly, the dry deposition velocities over water differ. From measurements over the ocean a mean value of $\overline{v}^\chem{O_3}_\text{water} = 0.019\,\unit{cm\,s^{-1}}$ has been deduced \citep{JGR:Helmig2012}. In a model study of different mechanisms of dry deposition to ocean waters by means of prescribed $\overline{v}_\text{water}^\chem{O_3}$ and one- and two-layer gas exchange modeling, \citet{ACP:Luhar2017} found $\overline{v}_\text{water}^\chem{O_3}$ ranging between $0.018\,\unit{cm\,s^{-1}}$ (two-layer scheme) and $0.039\,\unit{cm\,s^{-1}}$ (prescribed). With $\overline{v}^\chem{O_3}_\text{water} = (0.038\pm0.002)\,\unit{cm\,s^{-1}}$, the EMEP scheme (Section~\ref{subsubsec:Rb}) yields probably a too strong dry deposition to ocean which implicates that ozone concentrations might even become larger and dry deposition even lower in the model if a two-layer scheme would be implemented. With respect to forests, we might be able to improve the model performance further by allowing more PFTs, instead of using a broad average. 

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Comparison with observations}
\label{subsec:obs}
In this section, we compare our results to observations. For all comparisons with observations, we use the original resolution of the OsloCTM3 instead of the re-gridded.

In Fig.~\ref{fig:mmm_drydep_stations}a), seasonal cycles of average ozone dry deposition fluxes for six observation sites are shown. We have computed a model average for all sensitivity studies at the closest grid point and show the $1\,\sigma$ error band. The shaded area around the multi-model-mean indicates the broad range of model results but is not an actual error band since such is not given in \citet{ACP:Hardacre2015}. Although it is quite interesting to compare to the multi-model-mean, our focus lies on the performance of the OsloCTM3 with respect to the observational averages. At about $4$ of $6$ sites, the EMEP scheme performs better than the Wesely scheme and better than or the same as the multi-model-mean. We use a $\chi^2$-test
\begin{equation}
  \chi^2 = \sum_{i=1}^{12}\frac{\left(\overline{\chem{O_3}}_{\text{DD,\,}i}^\text{sim}-\overline{\chem{O_3}}_{\text{DD,\,}i}^\text{obs}\right)^2}{\sigma_i^2},
\end{equation}
with an estimated standard deviation of observation $\sigma_i=1\,\unit{mmol\,m^{-2}\,s^{-1}}$ and divide it by the number degrees of freedom (NDF) to assess this subjective analysis in a more objective way. The closer to $1$ this test scores, the better does the simulation represent the observation. A score between $0$ and $1$ indicates that the estimated $\sigma$ is too small. The results of the $\chi^2$-test are shown together with the divergences in Fig.~\ref{fig:mmm_drydep_stations}. The $\chi^2$-test reveals also that in $4$ of $6$ cases the EMEP scheme improves the performance of the OsloCTM3 with respect to ozone dry deposition fluxes, although a satisfying result is only achieved for two sites (Castel Porziano, Blodgett Forest). In the case of Castel Porziano, Auchencoth Moss, and Citrus Orchard, there may be a hint to an underlying model bias of a too small dry deposition of about $2.8\,\unit{mmol\,m^{-2}\,s^{-1}}$. With only one full year of simulation, the model uncertainty regarding the seasonal cycle at observational sites cannot be quantitatively assessed. Further more, the observational averages uses at most nine years worth of data which questions the statistical reliability of these data. Among other aspects that might affect the model performance the horizontal as well vertical resolution play an important role. In the rather coarse resolution, good matches between observation and model are only to be expected if the station is at a location which is representative for an area similar to the model resolution.

{\bf Additional zonal data and tropospheric ozone burden from satellites and sondes (\url{http://bodekerscientific.com/})?}

@@

{\bf Go through the finished text and resume/continue the started discussions in the evaluation section.}
dry dep velocity over water \citep{JGR:Helmig2012} Impact of ocean? Although it is a low number there is $\frac{2}{3}$ ocean on earth... (Think, I read this in Hardacre paper... -> \citet{ACP:Luhar2017}). Well, most of it is already answered in the last part of the model to model evaluation.
\citep{AE:Derwent2018} Uncertainties in surface ozone sources and sinks larger than in formulations? -> Need to read this one...

Come back to the discussion of the multi-model-mean - there seems to be a tendency that the ozone dry deposition is much lower in more advanced, newer models than in the old ones. May also include \citep{AE:Hu2017} in this part of the discussion.
%\subsection{HEADING}
%\subsubsection{HEADING}

\conclusions  %% \conclusions[modified heading if necessary]
\label{sec:conc}
What did we gain?

@@

\codedataavailability{TEXT} %% use this section when having data sets and software code available

@@

\authorcontribution{Stefanie Falk wrote the paper, did the data analysis, and finalized the implementation of the stomatal conductance in the dry deposition scheme. Amund S{\o}vde Haslerud did most of the implementation and documentation of the updated dry deposition scheme.} %% it is strongly recommended to make use of this section

>>>>

In Section~\ref{sec:model_des}, we will briefly describe the OsloCTM3, give a detailed account of the new dry deposition scheme (Section~\ref{subsec:DryDep}) as well as present pre-processing of meteorological input data to compute necessary input to the dry deposition scheme such as begin and duration of greening season (GDAY, GLEN) and photosynthetic photon flux density (PPFD) (Section~\ref{subsec:pre-pro}). In Section~\ref{sec:eval}, we present sensitivity tests with respect to a manifold of parameters in the dry deposition scheme (Section~\ref{subsec:sens}) and validate our results with respect to results from the multi-model intercomparison of \citet{ACP:Hardacre2015} (Section~\ref{subsec:model}) and to observations both at the surface as well as in the free troposphere (Section~\ref{subsec:obs}). In Section~\ref{sec:disc}, we will summarize our results and draw conclusions for further development of the model (Section~\ref{sec:conc}).
{\bf TODO: Precise outlook on paper when writing is done.}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Model description}
\label{sec:model_des}
The OsloCTM3 is a three dimensional global chemistry transport model (CTM). The key components of the OsloCTM3 have been described and evaluated by \citet{GMD:Sovde2012}. A detailed account of the capabilities of the OsloCTM3 in simulating anthropogenic aerosol forcing in the past and recent past using the Community Emission Data System (CEDS) historical emission inventory \citep{GMD:Hoesly2018} is given by \citet{GMD:Lund2018}. The OsloCTM3 can also be coupled to the Model of Emissions of Gases and Aerosols from Nature (MEGAN v2.10) \citep{ACP:Guenther2006}. A publication focusing on this is planed. Parallelization of the OsloCTM3 has not been optimized as of now. While the meteorological data driving the OsloCTM3 is given in a resolution of at T159N80L60, with the highest model level at $0.1\,\unit{hPa}$, it is very time and memory consuming to run the CTM with full chemistry at this resolution. Therefore, we run with a reduced horizontal resolution of $2.25^\circ\times2.25^\circ$. In the following, we will give a detailed account of the new dry deposition scheme. Although some of the equations in this section may be \emph{textbook knowledge}, for the sake of completeness, it is important to summarize them, nonetheless.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Ozone dry deposition scheme}
\label{subsec:DryDep}
In the original dry deposition scheme of the OsloCTM3, the state of the atmosphere was not taken into account. Dry deposition velocities were rather parameterized following the work of \citet{AE:Wesely1989} with parameter updates from \citet{JGR:Hough1991}. This means that seasonal day and night average deposition velocities for different land surface types (water, forest, grass, tundra/desert, and ice and snow) were in use. Day was distinguished from night by solar zenith angles below $90\,\unit{^\circ}$. Winter was defined by temperatures below $273.15\,\unit{K}$ for gridboxes containing land masses. For ocean, winter and summer parameters are equal in this parameterization, therefore no distinctive treatment is needed for ocean gridboxes. In addition, a reduced uptake due to snow cover above $1\,\unit{m}$ for forest and $10\,\unit{cm}$ for grass/tundra, respectively, was taken into account. We will refer to this parameterization as \emph{Wesely scheme}.\\

@@

\subsubsection*{Mapping of land surface types}
Land surface types from, e.g., \href{https://daac.ornl.gov/cgi-bin/dataset_lister.pl?p=29}{ISLSCP2} product from MODIS or Community Land Model (\href{http://www.cgd.ucar.edu/tss/clm/}{CLM}) 2 categories, have to be mapped to the nine land surface types used in the EMEP scheme (Fig.~\ref{fig:pft_mapping}). For both, MODIS and CLM~2 land surface categories, snow and ice cover is estimated from input meteorology, while water is defined as $1-\sum_{N=0}^{N_\text{max}} f_L(N)$. From the MODIS category \emph{Barren or sparsely vegetated}, everything polward from $60\,\unit{^\circ}$ is defined as tundra, while everything equatorward is categorized as desert. This mapping differs from the one used in the Wesely scheme.

@@

  \caption{Mapping of land surface categories from ISLSCP2 product of MODIS /  Community Land Model (CLM) 2 to the nine land surface types used in the EMAP scheme. Water bodies and snow and ice categories of MODIS are actually not mapped. For both, MODIS and CLM~2 land surface categories, snow and ice cover is estimated from input meteorology, while water is defined as $1-\sum_{N=0}^{N_\text{max}} f_L(N)$. From MODIS category 'Barren or sparsely vegetated', everything polward from $60\,\unit{^\circ}$ is defined as tundra, while everything equatorward is categorized as desert.}

@@

The annual total ozone dry deposition is shown in Fig.~\ref{fig:mmm_drydep}c). In accordance to the previously described features, we observe a reduction of the global total ozone dry deposition in all sensitivity studies to almost one half of the amount given by \emph{Wesely\_type}. The occurrence of this reduction in the zonal bands, where the major deserts are located, points to a change in $v_\text{desert}^\chem{O_3}$. Consulting the parameter file used in the Wesely scheme, we indeed find $v_\text{desert}^\chem{O_3} \equiv v_\text{tundra}^\chem{O_3} = 0.26\,\unit{cm\,s^{-1}}$ \citep{JGR:Hough1991}, while in the EMEP scheme $v_\text{desert}^\chem{O_3}=0.05\,\unit{cm\,s^{-1}}$ and $v^\chem{O_3}_\text{tundra} = 0.24\,\unit{cm\,s^{-1}}$, respectively. Similarly, dry deposition velocities over ice and snow and ocean have been even higher in the Wesely scheme ($v_\text{ice/snow}^\chem{O_3}\equiv v_\text{water}^\chem{O_3}=0.07\,\unit{cm\,s^{-1}}$) than in the original EMEP scheme ($v_\text{ice/snow}^\chem{O_3}\equiv v_\text{water}^\chem{O_3}=0.05\,\unit{cm\,s^{-1}}$). These differences in baseline dry deposition velocities over huge parts of the unvegetated surface of the Earth amount to most of the quantitative difference between the Wesely and the EMEP scheme. We further elaborate on this in the following section (Section~\ref{subsubsec:seasons}).

\begin{figure}[t]
  \includegraphics[width=8.3cm]{pictures/final-vmr_surfaceozone}\\
  \includegraphics[width=8.3cm]{pictures/final-total_ozone_drydepvelo}\\
  \includegraphics[width=8.3cm]{pictures/final-total_ozone_drydep}
  \caption{Comparison of the manifold of OsloCTM3 integrations with respect to (a) Surface ozone concentrations, (b) Annual average ozone dry deposition velocity, (c) Total annual ozone dry deposition. The different colors indicate sets of simulation with similar baselines. The multi-model mean from the evaluation of HTAP models by \citet{ACP:Hardacre2015} is shown as a reference (where available).}
  \label{fig:mmm_drydep}
\end{figure}
%
In Fig.~\ref{fig:mmm_drydep_season}a), the average zonal ozone dry deposition is shown separated by month. Where available, we have added the multi-model-mean given by \citet{ACP:Hardacre2015} as reference. There are two major phases apparent: NH and SH greening season. Former is reflected in a pronounced peak of $v_\text{DD}^\chem{O_3}$ in the northern mid latitudes, while the absence of this peak can be attributed to the latter in addition to a southward shift of the tropical peak dry deposition velocity. In the Wesely scheme, NH mid latitude peak velocities appear in June compared to July in the EMEP scheme, indicating that the seasonal cycles differ. The corresponding total monthly ozone dry deposition is shown in Fig.~\ref{fig:mmm_drydep_season}b). In general, the seasonal patterns are quite similar in the Wesely scheme and the EMEP scheme, displaying a strong symmetry around $10\,\unit{^\circ N}$ in January/February and November/December, respectively. What differs most is the molding and intensity of the NH peak dry deposition. Both schemes reach the maximum in June/July but the peak is much more differentiated in March already in the Wesely scheme. Similarly, the SH tropical peak dry deposition is reached in August/September but sustained longer, into October, in the Wesely scheme. Since we have not conducted any simulation with a meteorological year other than 2005, we cannot elaborate on whether this is a special feature of our chosen year or not. In comparison, the available two month of the multi-model-mean display a different pattern. Especially the NH/SH asymmetry is much more pronounced in February, than it is the EMEP scheme.
%
\begin{figure}[t]
  \includegraphics[width=8.3cm]{pictures/final-total_ozone_drydepvelo_monthly}\\
  \includegraphics[width=8.3cm]{pictures/final-total_ozone_drydep_monthly}
  \caption{Comparison of the manifold of OsloCTM3 integrations with respect to (a) Zonal average ozone dry deposition velocities; (b) Total annual amount of ozone removed from the atmosphere via dry deposition. The multi-model mean from the evaluation of HTAP models by \citet{ACP:Hardacre2015} is shown as a reference (where available).}
  \label{fig:mmm_drydep_season}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Average seasonal cycles}
\label{subsubsec:seasons}
%
To further disentangle the contributions of different regions to the global ozone budget, we will briefly look at different projections of seasonal cycles.

@@

As suggested by \citet{ACP:Hardacre2015}, we retrospectively separate ozone dry deposition velocities with respect to surface types. Based on a CLM~2 average dynamic land surface map, we generated masks for 9 different surface types (Fig.~\ref{fig:pft_landsurface}a) and used these to select gridboxes with a high percentage of these surface types, ranging from a meager $70\,\unit{\%}$ for cropland in the NH mid latitudes to $100\,\unit{\%}$ regarding desert, ocean, snow and ice, and tropical forest. Thus, it was not possible to exclusively select gridboxes with $100\,\unit{\%}$ cover for each surface type. Since we have not performed a full unfolding on the data, the results should be treated with slight caution. In Fig.~\ref{fig:mmm_drydep_season_pft}, the seasonal cycles of dry deposition velocities are shown for the nine surface categories. The patterns and absolute numbers differ substantially between the Wesely scheme and the EMEP scheme and the multi-model-mean. The divergence of the average dry deposition velocities between the Wesely scheme and the EMEP scheme in desert regions ($\Delta \overline{v}_\text{desert}^\chem{O_3} = 0.14\,\unit{cm\,s^{-1}}$) as well as grassland ($\Delta \overline{v}_\text{grassland}^\chem{O_3} = 0.4\,\unit{cm\,s^{-1}}$) is quite remarkable as well as the difference of both to the multi-model-mean in tropical forest regions ($\Delta \overline{v}_\text{tropical forest}^\chem{O_3} = 0.38\,\unit{cm\,s^{-1}}$). In contrast to the Wesely scheme and the EMEP scheme, the multi-model-mean indicates a rather pronounced seasonal cycle in desert regions ($0.10\,\unit{cm\,s^{-1}} \leq v^\chem{O_3}_\text{desert}\leq 0.15\,\unit{cm\,s^{-1}}$). The estimated dry deposition velocities over desert regions are consistent with the previously mentioned baseline values of ozone dry deposition, which means that in the EMEP scheme they are about one order of magnitude lower than in the Wesely scheme. From measurements of ozone fluxes in the Sahara desert, \citet{AE:Gusten1995} deduced $v^\chem{O_3}_\text{desert, day} = 0.1\,\unit{cm\,s^{-1}}$, $v^\chem{O_3}_\text{desert, night} = 0.04\,\unit{cm\,s^{-1}}$, and $\overline{v}^\chem{O_3}_\text{desert} = 0.065\,\unit{cm\,s^{-1}}$. This implies, that ozone dry deposition over desert regions is highly overestimated in the Wesely scheme as well as in the multi-model-mean, while it may be underestimated in the EMEP scheme. Similarly, the dry deposition velocities over water differ. From measurements over the ocean a mean value of $\overline{v}^\chem{O_3}_\text{water} = 0.019\,\unit{cm\,s^{-1}}$ has been deduced \citep{JGR:Helmig2012}. In a model study of different mechanisms of dry deposition to ocean waters by means of prescribed $\overline{v}_\text{water}^\chem{O_3}$ and one- and two-layer gas exchange modeling, \citet{ACP:Luhar2017} found $\overline{v}_\text{water}^\chem{O_3}$ ranging between $0.018\,\unit{cm\,s^{-1}}$ (two-layer scheme) and $0.039\,\unit{cm\,s^{-1}}$ (prescribed). With $\overline{v}^\chem{O_3}_\text{water} = (0.038\pm0.002)\,\unit{cm\,s^{-1}}$, the EMEP scheme (Section~\ref{subsubsec:Rb}) yields probably a too strong dry deposition to ocean which implicates that ozone concentrations might even become larger and dry deposition even lower in the model if a two-layer scheme would be implemented. With respect to forests, we might be able to improve the model performance further by allowing more PFTs, instead of using a broad average. 

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Comparison with observations}
\label{subsec:obs}
In this section, we compare our results to observations. For all comparisons, we use the original resolution of the OsloCTM3 instead of the re-gridded.

In Fig.~\ref{fig:mmm_drydep_stations}a), seasonal cycles of average ozone dry deposition fluxes for six observation sites are shown. We have computed a model average for all sensitivity studies at the closest grid point and show the $1\,\sigma$ error band. The shaded area around the multi-model-mean indicates the broad range of model results but is not an actual error band since such is not given in \citet{ACP:Hardacre2015}. At about $4$ of $6$ sites, the EMEP scheme performs better than the Wesely scheme and better than or the same as the multi-model-mean. We use a $\chi^2$-test
\begin{equation}
  \chi^2 = \sum_{i=1}^{12}\frac{\left(\overline{\chem{O_3}}_{\text{DD,\,}i}^\text{sim}-\overline{\chem{O_3}}_{\text{DD,\,}i}^\text{obs}\right)^2}{\sigma_i^2},
\end{equation}
with an estimated standard deviation of observation $\sigma_i=1\,\unit{mmol\,m^{-2}\,s^{-1}}$ and divide it by the number degrees of freedom (NDF) to assess this subjective analysis in a more objective way. The closer to $1$ this test scores, the better does the simulation represent the observation. A score between $0$ and $1$ indicates that the estimated $\sigma$ is too small. The results of the $\chi^2$-test are shown together with the divergences in Fig.~\ref{fig:mmm_drydep_stations}. The $\chi^2$-test reveals also that in $4$ of $6$ cases the EMEP scheme improves the performance of the OsloCTM3 with respect to ozone dry deposition fluxes, although a satisfying result is only achieved for two sites (Castel Porziano, Blodgett Forest). In the case of Castel Porziano, Auchencoth Moss, and Citrus Orchard, there may be a hint to an underlying model bias of a too small dry deposition of about $2.8\,\unit{mmol\,m^{-2}\,s^{-1}}$. With only one full year of simulation, the model uncertainty regarding the seasonal cycle at observational sites cannot be quantitatively assessed. Further more, the observational averages uses at most nine years worth of data which questions the statistical reliability of these data. Among other aspects that might affect the model performance the horizontal as well vertical resolution play an important role. In the rather coarse resolution, good matches between observation and model are only to be expected if the station is at a location which is representative for an area similar to the model resolution.

{\bf TODO: Additional zonal data and tropospheric ozone burden from satellites and sondes (\url{http://bodekerscientific.com/})?}

@@

{\bf Go through the finished text and resume/continue the started discussions in the evaluation section.}
dry dep velocity over water \citep{JGR:Helmig2012} Impact of ocean? Although it is a low number there is $\frac{2}{3}$ ocean on earth... (Think, I read this in Hardacre paper... -> \citet{ACP:Luhar2017}). Well, most of it is already answered in the last part of the model to model evaluation.
\citep{AE:Derwent2018} Uncertainties in surface ozone sources and sinks larger than in formulations? -> Need to read this one...
Resolution and representativeness of site
Come back to the discussion of the multi-model-mean - there seems to be a tendency that the ozone dry deposition is much lower in more advanced, newer models than in the old ones. May also include \citep{AE:Hu2017} in this part of the discussion.
%\subsection{HEADING}
%\subsubsection{HEADING}

\conclusions  %% \conclusions[modified heading if necessary]
\label{sec:conc}
We have presented an update of the dry deposition scheme in the OsloCTM3 from a purely prescribed parameterization based on the works of \citet{AE:Wesely1989,JGR:Hough1991} to a more process oriented parameterization taking the state of the atmosphere into account \citep{ACP:Simpson2012}. In general, this update decreases the annual amount of ozone dry deposition up to $100\,\unit{\%}$ especially over large desert regions and the oceans. Most of this decrease is due to changes in the baseline dry deposition velocities at the order of one magnitude in case of desert and grasslands or the accumulative effect of small changes over the ocean in the southern hemisphere. With respect to available measurements of dry deposition velocities of ozone over desert \citep{AE:Gusten1995} and ocean \citep{JGR:Helmig2012}, the OsloCTM3 may slightly underestimate ozone dry deposition in the former case and overestimate it in the latter. Regarding the vastness of the ocean and the ongoing desertification, it may be worthwhile studying dry deposition in these regimes at a more process-based level (e.g., 2-layer gas exchange with ocean waters). Although ice and snow only amounts to less than $1\,\unit{\%}$ to the global annual ozone dry deposition, a change in dry deposition velocities not only almost doubles the surface ozone concentrations in the high Arctics but also affects surface ozone down to latitudes at about $50\,\unit{^\circ}$. In comparison to multi-model-mean results, we conclude that it is important to use an updated ozone dry deposition to ice and snow \citep{ACP:Helmig2007}. We have extensively studied the parameter space of the stomatal conductance parameterization and found the largest sensitivity of total surface ozone in the tropics. This may indicate that changes in solar radiation due to changes cloud and aerosol may also strongly influence surface ozone in the tropics. In the most extreme test scenario, allowing for an unlimited photosynthesis, the increase in surface ozone amounts to up to $16\,\unit{\%}$ compared to the reference experiment. Switching off the phenology in the tropics only increases ozone there by up to $10\,\unit{\%}$, while the complete switch-off amounts only to a maximum increase of $5\,\unit{\%}$. Total surface ozone increases by about $7.7\,\unit{\%}$ in the northern hemisphere ($(50-75)\,\unit{^\circ N}$) if a greening season based on a $5\,\unit{^\circ C}$-days criteria is used. An important factor in the global ozone budget are also emissions of precursor substances. In a comparison between 2005 and 2014 emissions using the same meteorology, we found a decrease in total surface ozone up to $5\,\unit{\%}$ in the tropics. We also evaluated the model with observed dry deposition velocities at six sites in the northern hemisphere and found that the updated dry deposition scheme performs better than the old one, but is not able to reproduce the measurements at most sites well. This may be due to several reasons. The rather low model resolution in both horizontal and vertical affects the distribution and transport (e.g., long-range, stratosphere--troposphere exchange) as well as the regional production of ozone. Depending on the location of the observation site and its respective representativeness for a larger area ozone dry deposition and ozone concentrations will be over- or underestimated. In addition, a comparison of very few years of measurement to only one specific year of simulation may reflect the year to year variability more than the actual model performance.
For further diagnostics, we shall make dry deposition velocities directly available in the model output.

@@

\codedataavailability{The OsloCTM3 is made publicly available on git-hub under a MIT license. Model results can be made available under request.} %% use this section when having data sets and software code available

@@

\authorcontribution{Stefanie Falk has finalized the implementation of the stomatal conductance in the EMEP-based dry deposition scheme of the OsloCTM3, conducted the tests, and evaluated the results. Amund S{\o}vde Haslerud has implemented the EMEP-based dry deposition scheme and wrote the documentation. Both authors contributed to the writing of the paper.} %% it is strongly recommended to make use of this section
